---
title: "Basic functions for retreaving results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(vmiBiomass)
library(dplyr)
library(ggplot2)

```

# Introduction

The package contains the estimated parameters, and their uncertainties, for the 36 curve estimates for the paper "Age-dependence of stand biomass in managed boreal forests based on the Finnish National Forest Inventory data" by Repo et al 2021.

The idea is to connect biomass of a forest plot with its (estimated) age. A parametric function is chosen to model the connection. Based on initial screening, we chose the Gompertz-function, using the parametrisation

$$f(x) = a + A\exp[-mk^x]$$

- $a$: Nugget effect
- $A$: upper asymptote
- $m$: lag parameter
- $k$: growth rate


The response variable for each group $g=1,...,36$ is biomass $y_{si}$, grouped within stands $s$. 

For each group, four versions were fitted. 

1. Additive model "A": $$y_{si}\sim N(f(age_{si}), \sigma_{e}^2)$$
2. Additive model with stand-wise random effect "A + re": $$y_{si}\sim N(u_s +f(age_{si}), \sigma_{e}^2)$$ with $u_s\sim N(0, \sigma_s^2)$
3. Multiplicative model "M": $$\log(y_{si})\sim N(\log[f(age_{si})], \sigma_{e}^2)$$
4. Multiplicative model with stand-wise random effect "M + re": $$\log(y_{si})\sim N(\log[f(age_{si})] + u_s, \sigma_{e}^2)$$ with $u_s\sim N(0, \sigma_s^2)$


All results are included in this package. The priors are available in the code, see next section.

# The Stan-model code
The models were fitted using the package `rstan`. The model code is available in text-objects, e.g.

```{r}
cat(mMre.stan)
```

For more information, see `?mA.stan`.


# Retrieving parameter estimates

The posterior estimates for the parameters results are stored in named-list. The group-information is stored in the object `lookup` 

```{r}
head(lookup)
```
from which the package-wide group-identifier `id` can be read for any group and used for retrieving specific results. For example,

```{r}
l1 <- lookup %>% filter(region == "SOUTH" & soil == "mineral" & fertility == "4" &
                    err_model == "M" &with_ranefs == TRUE)
l1
```

Then we can use

```{r}
r1 <- getParameters(id = l1$id)
head(r1)
```

A shortcut is 

```{r}
r1 <- getParameters(region = "SOUTH", soil = "mineral", fertility = "4", err_model = "M", with_re = TRUE)
```

See the documentation for `getResults` for more information.

# Retrieving curve estimates

To get posterior estimates of the curves on the range $[0,200]y$:

```{r}
c1 <- getCurves(region = "SOUTH", soil = "mineral", fertility = "4", with_re = TRUE, err_model = c("A", "M"))
```
The age is in centuries.
```{r, fig.width=9}
c1 %>% ggplot(aes(x * 100, mean)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`, fill = err_model), alpha = .5) +
  geom_line(aes(col = err_model))
```

Note that due to the non-linear nature of the curve, there is a slight discrepancy between the posterior curves and the curve at the posterior parameters:

```{r, fig.width=9}
c1_1 <- getCurves(id = l1$id, log_scale = TRUE)
c1_2 <- evalGompertz(id = l1$id, log_scale = TRUE, x = c1_1$x)

bind_rows(c1_1 %>% mutate(m = "posterior mean curve"),
          c1_2 %>% mutate(m = "curve at posterior mean")) %>%
  ggplot( aes(x, mean, col = m)) + 
  geom_line()

```

# Retrieving everything

All stored information, at the moment undocumented, is available in the `results`-object, for example the posterior covariance of parameter estimates:

```{r}
res1 <- results[[ l1$id ]]
res1$est$Cov
```

```{r}
str(res1, 1)
```

